Hierarchically modelling \emph{Kepler} dwarfs and subgiants to improve inference of stellar properties with asteroseismology

### 2020-11-17 paper/main.tex
More typo fixes
Updated tables

% Abstract of the paper <= 250 words
\begin{abstract}
    With recent advances in modelling stars using high-precision asteroseismology, the systematic effects associated with our assumptions of stellar helium abundance ($Y$) and the mixing-length theory parameter ($\mlt$) are becoming more important. We apply a new method to improve the inference of stellar parameters for a sample of \emph{Kepler} dwarfs and subgiants across a narrow mass range ($0.8 < M < \SI{1.2}{\solarmass}$). In this method, we include a statistical treatment of $Y$ and the $\mlt$. We develop a hierarchical Bayesian model to encode information about the distribution of $Y$ and $\mlt$ in the population, fitting a linear helium enrichment law including an intrinsic spread around this relation and normal distribution in $\mlt$. We test various levels of pooling parameters, with and without solar data as a calibrator. When including solar data, we find the gradient for the enrichment law, $\Delta Y / \Delta Z = 1.05\substack{+0.28\\-0.25}$ and the mean $\mlt$ in the population, $\mu_\alpha = 1.90\substack{+0.10\\-0.09}$. While accounting for the uncertainty in $Y$ and $\mlt$, we are still able to report statistical uncertainties of 2.5 per cent in mass, 1.2 per cent in radius, and 12 per cent in age. Our method can also be applied to larger samples which will lead to improved constraints on both the population level inference and the star-by-star fundamental parameters. We show that more work is needed to study the validity of using the Sun as a calibrator for $Y$ and $\mlt$, perhaps by including asteroseismic signatures of helium abundance.
\end{abstract}

@@

We chose to ignore the well known, yet poorly characterised impact of modelled oscillation mode inaccuracies in the near-surface region of the star \citep{Kjeldsen.Bedding.ea2008, Ball.Gizon2014, Sonoi.Samadi.ea2015}. This typically presents only a small effect compared to observational uncertainties when considering the average large frequency spacing, $\dnu$. A thorough treatment of systematics related to $\dnu$ is instead left to future work. % This may need more

@@

\subsection{Solar calibrator}\label{sec:sun}

@@

Pooling parameters in an HBM allows us to use the Sun as a calibrator in a unique way. Rather than calibrating our model physics to the Sun and then assuming the calibrated parameters across our sample, we can include the Sun as a part of the same population. In other words, if we assume $Y_\mathrm{init}$ and $\mlt$ for the Sun are a part of the same prior distribution as for the sample, then we can add Solar data to our model inputs.

For both the PP and MP models, we ran versions with and without data for the Sun included in the APOKASC population, referred to as PPS and MPS respectively. We adopted standard solar data as shown in Table \ref{tab:sun} with uncertainties conservatively limited to the accuracy of the ANN for $R$, $L$ and representative of variation in the literature for $\teff$. We also adopted $\dnu=\SI{135.1(2)}{\micro\hertz}$ from \citet{Huber.Bedding.ea2011} with a standard deviation representative of variations in measurements of the solar $\dnu$ \citep{Broomhall.Chaplin.ea2011}.

\begin{table}
    \centering
    \caption{Solar input data where $\mu$ and $\sigma$ are the means and standard deviations respectively.}
    \label{tab:sun}
    \input{sun_inputs.tex}
\end{table}

@@

@@

During an initial run of the NP model, we dropped 16 of the 81 stars from the sample. Of the removed stars, we found the posteriors in $M$ for 6 skewed towards the prior upper mass limit of \SI{1.2}{\solarmass}. The remaining 10 removed stars suffered poor convergence during sampling ($\hat{r} >> 1.04$) which could be because of poor step-size tuning and sampling at the prior boundary.

In Table \ref{tab:np}, we preset results for the remaining 65 stars from the NP model. Running the NP model with synthetic stars resulted in unreliable uncertainties (see Appendix \ref{sec:test-stars}). This was because the boundary of the priors in $Y_\mathrm{init}$ and $\mlt$ truncated the posterior distribution leading to underestimated uncertainties and skewing their posterior means towards the centre of their priors. Therefore, we present the NP results only for comparison purposes, but we exclude them from further discussion.

@@

Out of the 65 stars with results from the NP model, 2 stars were dropped from the PP model with problems during sampling. A consequence of partially pooling parameters is that a population spread, $\sigma$ allows for individual parameters to vary outside of the prior range given to the population mean, $\mu$. In this case, individual stellar $\mlt$ was allowed to vary outside the range for which the ANN was valid if $\sigma_\alpha$ was large. The 2 removed stars happened to have high likelihoods outside of the valid $\mlt$ range. We found that removing the same 2 stars from the other models made negligible difference to the results, so we leave a solution to this problem to future work. In Tables \ref{tab:pp} and \ref{tab:pps} we present the results for the remaining 63 stars from the PP and PPS model respectively.

When we max-pooled $\mlt$ we did not see the same issue with the 2 removed stars from the PP model because we did not include any population spread. Therefore, we kept these stars in the sample and present results for 65 stars in Tables \ref{tab:mp} and \ref{tab:mps} for the MP and MPS models respectively.

@@

We obtained values for the hyperparameters for each of the models and present them in Table \ref{tab:hparam_results} along with their upper and lower 68 per cent credible regions. We omitted the results for $Y_P$ because its posterior was the same as the prior, $Y_P=0.247\pm0.001$ for all the models. We fit the same hyperparameters from the PP model to the NP model results for $Y_\mathrm{init}$, $Z_\mathrm{init}$ and $\mlt$ for the purpose of comparison. However, the NP model suffered from boundary effects which made the resulting fit unreliable, pushing the population mean to the centre of the priors and underestimating the uncertainties. We leave the NP results here only for completeness.

@@

Figure \ref{fig:corners-pp} shows the joint and marginal distributions (corner plot) output by the PP and PPS model. We saw an anti-correlation between $\Delta Y / \Delta Z$ and $\mu_\alpha$, expected due to the degeneracy between the two parameters in the stellar evolutionary models. In Figure \ref{fig:corners-mp}, we also show the corner plot for the MP and MPS model output. Similarly, we see an anti-correlation between $\Delta Y / \Delta Z$ and $\mlt$.

@@

We present 100 random samples from the posterior for the helium enrichment relation from the PPS model in Figure \ref{fig:helium}. In this figure, we also plot the individual results for $Y_\mathrm{init}$ and $Z_\mathrm{init}$ for each of the stars in the sample. This is an example of \emph{shrinkage} in the HBM; the estimates for individual stellar parameters move towards the mean of the population.

@@

When we added the Sun to the pooled models, PPS and MPS, obtained $\Delta Y / \Delta Z$ up to 2-$\sigma$ lower than the models without the Sun. In both models, the resulting $\Delta Y / \Delta Z$ of approximately \numrange{0.8}{1.0} was consistent with the initial helium fraction expected from solar models with our choice of \citet{Asplund.Grevesse.ea2009} abundances \citep{Serenelli.Basu2010}. However, such solar models have been shown to not recover helioseismic measurements of helium in the Sun \citep{Basu.Antia2004, Serenelli.Basu.ea2009, Villante.Serenelli.ea2014}. Solar models with the older \citet{Grevesse.Sauval1998} abundances typically yield higher helium fractions more in-line with helioseismology. Interestingly, the $\Delta Y / \Delta Z$ from the PP and MP models are higher than those including the Sun, in closer agreement to those calibrated with the older abundances. We could extend our model to include asteroseismic indicators of helium to improve the uncertainties on $Y$ and test whether this difference becomes more significant.

@@

To a greater degree then chemical composition, the best-fitting $\mlt$ depends on the choice of model physics and stellar modelling code. The mixing-length theory is an approximation of convection which is often calibrated to the Sun and then assumed for all stars in a model. However, studies of 3D hydrodynamical simulations suggest that the degree in which $\mlt$ approximates convection varies across the HR diagram \citep{Magic.Weiss.ea2015} and this is confirmed when modelling stars with asteroseismology \citep{Tayar.Somers.ea2017}.

Our fit to the NP model results obtained a result for $\mu_\alpha$ close to $2.0$ -- the midpoint of its prior distribution. This was expected because the tests on synthetic stars showed that $\mlt$ was biased in this way due to boundary effects from truncating the distribution at $1.5$ and $2.5$. Again, the NP hyperparameter fit is used for comparison, but is ultimately unreliable.

@@

In the following subsections, we compare the results between our PPS model with that of \citetalias{Serenelli.Johnson.ea2017} with reference to Figure \ref{fig:comp}. In particular, we compare the mass, radius and age results and find good agreement overall.

@@

\subsubsection{Mass}

@@

\subsubsection{Radius}

@@

\subsubsection{Age}

@@

In previous work studying stars in the APOKASC sample, several pipelines used a range of stellar evolutionary codes and model physics are employed to evaluate systematic uncertainties from the models \citep{Serenelli.Johnson.ea2017, SilvaAguirre.Lund.ea2017}. Using a hierarchical model in this work enabled us to reduce median statistic uncertainties to 2.5 per cent in mass, 1.2 per cent in radius and 12 per cent in age. The systematic uncertainty analysis of \citetalias{Serenelli.Johnson.ea2017} found median systematics of 3, 1 and 13 per cent in mass, radius and age respectively. Reducing statistical uncertainties highlights the importance of understanding systematics.

@@

\subsection{The potential of pooling}\label{sec:future}

Including all-sky data from \emph{TESS} and in anticipation of \emph{Plato} \citep{Rauer.Catala.ea2014} we can expect our sample size of asteroseismic dwarfs and subgiants only to increase. There is also scope to extend our grid of models to include red giants, for which there are vast catalogues of stars already studied with \emph{Kepler} \citep{Pinsonneault.Elsworth.ea2018}.

@@

>>>>

% Abstract of the paper <= 250 words
\begin{abstract}
    With recent advances in modelling stars using high-precision asteroseismology, the systematic effects associated with our assumptions of stellar helium abundance ($Y$) and the mixing-length theory parameter ($\mlt$) are becoming more important. We apply a new method to improve the inference of stellar parameters for a sample of \emph{Kepler} dwarfs and subgiants across a narrow mass range ($0.8 < M < \SI{1.2}{\solarmass}$). In this method, we include a statistical treatment of $Y$ and the $\mlt$. We develop a hierarchical Bayesian model to encode information about the distribution of $Y$ and $\mlt$ in the population, fitting a linear helium enrichment law including an intrinsic spread around this relation and normal distribution in $\mlt$. We test various levels of pooling parameters, with and without solar data as a calibrator. When including the Sun as a star, we find the gradient for the enrichment law, $\Delta Y / \Delta Z = 1.05\substack{+0.28\\-0.25}$ and the mean $\mlt$ in the population, $\mu_\alpha = 1.90\substack{+0.10\\-0.09}$. While accounting for the uncertainty in $Y$ and $\mlt$, we are still able to report statistical uncertainties of 2.5 per cent in mass, 1.2 per cent in radius, and 12 per cent in age. Our method can also be applied to larger samples which will lead to improved constraints on both the population level inference and the star-by-star fundamental parameters. We show that more work is needed to study the validity of using the Sun as a calibrator for $Y$ and $\mlt$, perhaps by including asteroseismic signatures of helium abundance.
\end{abstract}

@@

We chose to ignore the well known, yet poorly characterised impact of modelled oscillation mode inaccuracies in the near-surface region of the star \citep{Kjeldsen.Bedding.ea2008, Ball.Gizon2014, Sonoi.Samadi.ea2015}. This typically presents only a small effect compared to observational uncertainties when considering the average large frequency spacing, $\dnu$. Additionally, there may be further inaccuracies in the modelled $\dnu$ because of variations in p mode frequencies with stellar activity \citep{Chaplin.Elsworth.ea2007, Kiefer.Schad.ea2017}. Therefore, a thorough treatment of systematic uncertainties in $\dnu$ is instead left to future work (Carboneau et al. in preparation). % This may need more

@@

\subsection{The Sun as a star}\label{sec:sun}

@@

Pooling parameters in an HBM allows us to use the Sun as a calibrator in a unique way. Rather than calibrating our model physics to the Sun and then assuming the calibrated parameters across our sample, we can include the Sun as a part of the same population as our sample of stars. If we assume $Y_\mathrm{init}$ and $\mlt$ for the Sun are a part of the same prior distribution as for the rest of the sample, then we can simply add solar observables to our model inputs.

For both the PP and MP models, we iterated with and without data for the Sun included in the population, referred to as PPS and MPS respectively. We adopted the solar data in Table \ref{tab:sun} with uncertainties conservatively limited to the accuracy of the ANN for $R$, $L$ and representative of variation in the literature for $\teff$. We also adopted $\dnu=\SI{135.1(2)}{\micro\hertz}$ with a central value from \citet{Huber.Bedding.ea2011} and a standard deviation representative of variations in measurements of the solar $\dnu$ \citep{Broomhall.Chaplin.ea2011}.

\begin{table}
    \centering
    \caption{Solar input data. The references correspond to the central values and the uncertainties are chosen to either be representative of the ANN accuracy or the spread of values in the literature (see text for details).}
    \label{tab:sun}
    \input{sun_inputs.tex}
\end{table}

@@

Since our initial sample was chosen based on masses from \citetalias{Serenelli.Johnson.ea2017}, we expected some stars to lie outside (or near the boundary) of the observational parameter space provided by our grid of stellar models. We used an initial run of the NP model to catch and remove these stars. During the initial run, we dropped 16 of the 81 stars from the sample. Of the removed stars, we found the posteriors in $M$ for 6 skewed towards the prior upper mass limit of \SI{1.2}{\solarmass}. The remaining 10 removed stars suffered poor convergence during sampling ($\hat{r} >> 1.04$) which could be because of poor step-size tuning and sampling at the prior boundary.

Out of the remaining 65 stars with results from the NP model, 2 stars were dropped from the PP model. A consequence of partially pooling parameters is that a population spread, $\sigma$ allows for individual parameters to vary outside of the prior range given to the population mean, $\mu$. In this case, individual stellar $\mlt$ was allowed to vary outside the range for which the ANN was valid if $\sigma_\alpha$ was large. The 2 removed stars happened to have high likelihoods outside of the valid $\mlt$ range. We found that removing the same 2 stars from the other models made negligible difference to the results, so we leave a solution to this problem to future work. Naturally, we did not see the same issue in the MP model, so we proceeded with modelling the same 65 stars as with the NP model.

@@

In Table \ref{tab:np}, we present results for the 65 APOKASC stars from the NP model. Running the NP model with synthetic stars resulted in unreliable uncertainties (see Appendix \ref{sec:test-stars}). This was because the boundary of the priors in $Y_\mathrm{init}$ and $\mlt$ truncated the posterior distribution leading to underestimated uncertainties and skewing their posterior means towards the centre of their priors. Therefore, we present the NP results only for comparison purposes, but we exclude them from further discussion. In Tables \ref{tab:pp} and \ref{tab:pps} we present the results for the 63 stars from the PP and PPS model respectively. In Tables \ref{tab:mp} and \ref{tab:mps} we also present results for the 65 stars from the MP and MPS models respectively. We note that for the MP models, there is no column for $\mlt$ because this parameter is the same across the population and hence is given in Section \ref{sec:hparam-results}.

@@

@@

We obtained values for the hyperparameters for each of the models and present them in Table \ref{tab:hparam_results} along with their upper and lower 68 per cent credible regions. We omit the results for $Y_P$ because its posterior is the same as the prior, $Y_P=0.247\pm0.001$ for all the models. We fit the same hyperparameters from the PP model to the NP model results for $Y_\mathrm{init}$, $Z_\mathrm{init}$ and $\mlt$ for the purpose of comparison. However, the NP model results suffer from boundary effects which makes the resulting fit unreliable, pushing the population mean to the centre of the priors and underestimating the uncertainties. We leave the NP results here only for completeness.

@@

Figure \ref{fig:corners-pp} shows the joint and marginal distributions (corner plot) output by the PP and PPS model. We see an anti-correlation between $\Delta Y / \Delta Z$ and $\mu_\alpha$, expected due to the degeneracy between the two parameters in the stellar evolutionary models. In Figure \ref{fig:corners-mp}, we also show the corner plot for the MP and MPS model output. Similarly, we see an anti-correlation between $\Delta Y / \Delta Z$ and $\mlt$.

@@

We present the helium enrichment relation resulting from the PPS model in Figure \ref{fig:helium}. In this figure, we plot the individual results for $Y_\mathrm{init}$ and $Z_\mathrm{init}$ for each of the stars from the NP and PPS models. This is an example of shrinkage in the HBM; the estimates for individual stellar parameters move towards the mean of the population when they are pooled.

@@

When we added the Sun to the pooled models, PPS and MPS, obtained $\Delta Y / \Delta Z$ up to 2-$\sigma$ lower than the models without the Sun. In both models, the resulting $\Delta Y / \Delta Z$ of approximately \numrange{0.8}{1.0} was consistent with the initial helium fraction expected from solar models with our choice of \citet{Asplund.Grevesse.ea2009} abundances \citep{Serenelli.Basu2010}. However, such solar models have been shown to not recover helioseismic measurements of helium in the Sun \citep{Basu.Antia2004, Serenelli.Basu.ea2009, Villante.Serenelli.ea2014}. Solar models with the older \citet{Grevesse.Sauval1998} abundances typically yield higher helium fractions more in-line with helioseismology. The $\Delta Y / \Delta Z$ from the PP and MP models are higher than those including the Sun, in closer agreement to those calibrated with the older abundances. We could extend our model to include asteroseismic indicators of helium to improve the uncertainties on $Y$ and test whether this difference becomes more significant.

@@

To a greater degree than chemical composition, the best-fitting $\mlt$ depends on the choice of model physics and stellar modelling code. The mixing-length theory is an approximation of convection which is often calibrated to the Sun and then assumed for all stars in a model. However, studies of 3D hydrodynamical simulations suggest that the degree to which $\mlt$ approximates convection varies across the HR diagram \citep{Magic.Weiss.ea2015} and this is confirmed when modelling stars with asteroseismology \citep{Tayar.Somers.ea2017}.

% Our fit to the NP model results obtained a result for $\mu_\alpha$ close to $2.0$ -- the midpoint of its prior distribution. This was expected because the tests on synthetic stars showed that $\mlt$ was biased in this way due to boundary effects from truncating the distribution at $1.5$ and $2.5$. Again, the NP hyperparameter fit is used for comparison, but is ultimately unreliable.

@@

In the following subsection, we compare the results between our PPS model with that of \citetalias{Serenelli.Johnson.ea2017} for mass, radius and age with reference to Figure \ref{fig:comp}. We preferred the PPS model for comparison because it utilised the high-precision data available for the Sun as a star to help calibrate the sample, while partially pooling both $Y_\mathrm{init}$ and $\mlt$ to allow for small variations within the population.

@@

\subsubsection{Mass, radius and age}

@@

@@

@@

In previous work studying stars in the APOKASC sample, several pipelines used a range of stellar evolutionary codes and model physics are employed to evaluate systematic uncertainties from the models \citep{Serenelli.Johnson.ea2017, SilvaAguirre.Lund.ea2017}. Using a hierarchical model in this work enabled us to reduce median statistic uncertainties to 2.5 per cent in mass, 1.2 per cent in radius and 12 per cent in age. The systematic uncertainty analysis of \citetalias{Serenelli.Johnson.ea2017} found median systematic uncertainties of 3, 1 and 13 per cent in mass, radius and age respectively. Reducing statistical uncertainties highlights the importance of understanding systematics uncertainties.

@@

% \subsection{The potential of pooling}\label{sec:future}

@@

Including all-sky data from \emph{TESS} and in anticipation of \emph{PLATO} \citep{Rauer.Catala.ea2014} we can expect our sample size of asteroseismic dwarfs and subgiants only to increase. There is also scope to extend our grid of models to include red giants, for which there are vast catalogues of stars already studied with \emph{Kepler} \citep{Pinsonneault.Elsworth.ea2018}.
